---
title: 'Geog 4/6300: Lab 3'
output:
  github_document: default
---

```{r setup, include=FALSE}
#install.packages("sf")
#install.packages("tmap")
#install.packages("spatstat")

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(tmap)
library(spatstat)
library(knitr)
```

## Spatial statistics and visualization

**Value:** 30 points

**Overview:**
This lab has three main sections: mapping point patterns, calculating rates and location quotients, and quadrat analysis. Your answers to the lab questions should be typed in the relevant spaces below and pushed to your Github repo. 

###Part 1: mapping the data and its distribution###
The data folder has two files related to crime data for Spokane, Washington.  

* “Spokane_crimes_all” has data for every crime committed in from 2000-2015. 

* "Spokane_precincts" has precinct boundaries for the city of Spokane and its population 2015. This is saved as a geojson file.

In the first section of this lab, you'll load this data and analyze it using several metrics related to point pattern analysis.

**Question 1 (3 points)** _Load the crime reports (Spokane_crimes_all) file using read_csv. Then use st_read to import the precinct boundaries. Call the head of the Spokane_crimes_all file using kable when you are done._

```{r}
spokane_crimes_all <- read_csv("data/Spokane_crimes_all.csv")
precincts_spok <- st_read("data/Spokane_precincts.geojson")

kable(head(spokane_crimes_all))
```

**Question 2 (3 points):** _Filter the data to include just arson, assault, and vehicle theft. Use group_by and summarise to calculate the mean centers for each type of offense. Then use ggplot and geom_point to plot these points on a graph, changing the color based on the offense._

```{r}
AR_AS_VT_rate<- spokane_crimes_all%>%
  group_by(Offense)%>%
   filter(Offense == "Arson" | Offense == "Assault" | Offense == "Vehicle Theft")%>%
  
# Use long and lat data to calculate the mean center:
# This is done by taking the sum of all the latitude values and then dividing by the count of the latitude column. The same thing is repeated in the Y direction - All values in the longitude column are summed and then divided by the count of observations in that column. The values are recorded and saved as two SEPERATE and new variable. Once those variables have been created, the two columns can be joined together to form a final variable that contains the the specific coordinate information.
  
  summarise(mean_cent_X = mean(Long), 
            mean_cent_Y = mean(Lat))

ggplot(data = AR_AS_VT_rate, aes(x = mean_cent_X, y = mean_cent_Y, color = Offense))+
  geom_point()+
  xlab("Longitude")+
  ylab("Latitude")
```

**Question 3 (2 points)** _Interpret how geographic differences between the mean centers you calculated are meaningful. What do they tell you about the locations of these crimes?_

{The difference in mean center is directly affected by the spatial components used in its calculation. Due to that fact, the disparity in mean center between the three variables may be indicative of a overarching geographic factor. An example of this might be with regards to the vehicle theft mean center. That mean center may be located where it is because the geographic area that surrounds may be particularly suceptible to car thefts because it's a high income area, potentially. This data gives us an idea that particular crimes or types of crimes may traceback to particular geographic locations in higher proportions when compared to the number of equivalent crimes in a different geographic area.}

**Question 4 (3 points):** _Filter the crime reports so that it contains just murder as the offense type, and then use st_as_sf to convert these to spatial data. Using tmap, create a map that shows the location of all reports in the this data frame and the precinct boundaries. Also include a scale bar and north arrow. See class scripts for a model to work from._

```{r}
murder_reports <- spokane_crimes_all%>%
  filter(Offense == "Murder")
murder_reports_sf <- st_as_sf(murder_reports,
                              coords=c(x = "Long", y = "Lat"),#indicate coordinate columns
                              crs=4326, #Set the projections
                              remove=FALSE) #Keep the coordinate columns


pre_murd_join<-murder_reports_sf%>%
  st_join(precincts_spok, join = st_within)

#my_bbox<-c(xmin=min(murder_reports_sf$Long),
#           ymin=min(murder_reports_sf$Lat),
#           xmax=max(murder_reports_sf$Long),
#           ymax=max(murder_reports_sf$Lat))

#murder_rep_map <- precincts_spok%>%
#  st_crs(murder_rep_map) = 4326
#  st_join(murder_reports_sf)

tmap_mode("plot")

tm_shape(precincts_spok)+
  tm_polygons()+
  tm_shape(pre_murd_join)+
  tm_dots(size = 0.1, alpha = 0.5, col = "red")+
tm_compass(position=c("right", "top"))+
tm_scale_bar(position="left")+
tm_legend(legend.outside=TRUE)

#ggplot(pre_murd_join)+
 # geom_sf()


```

**Question 5 (2 points):** _The st_as_sf function has multiple parameters (such as `coords`). Briefly explain the purpose of each one you used for question 4._

{The first argument included inside of the st_as_sf function specifies the data set to use. The second argument establishes where the coordinates needed to construct the right geometry are supposed to come from - in this case, the Longitude column from the df in question [murder_reports] represents the x value data and the latitude corresponds to what will become the y values. The third argument of the st_as_sf function "crs" details what map projection when constructing the map. The final argument details whether to keep the columns detailing the original coordinate information or not; "TRUE" causes the original columns to be removed since their geometry has been reported in a new column. If remove=FALSE, then the original columns that contained the coordinate information will me kept.}

###Part 2: Crime rates###
In the second section of this lab, you'll do a spatial join of points to precincts to identify the precinct ID of each reported crime. You'll then summarise these reports to create a crime rate for each precinct, which is usually reported as crime per 100,000 people.

**Question 6 (3 points):** _Link the crimes from 2012 to the precinct they were in. To do this, first filter the points based on year and then match points to their precinct using a spatial join. Use st_as_sf to convert the crime reports to spatial data. Then use st_join to link precinct information to the crime reports. When that's done, remove the geometry column using st_set_geometry and then filter out any points that do not have a matching precinct. Call the head of this table using kable when you are done._

```{r}
crimes_2012<-spokane_crimes_all%>%
  filter(Year=="2012")%>%
  st_as_sf(coords=c(x = "Long", y = "Lat"), 
           crs=4326,
           remove=TRUE)

crimes_2012_joint<-crimes_2012%>%
  st_join(precincts_spok)%>%
  st_set_geometry(NULL)%>%
  na.omit(crimes_2012_joint)
  
kable(head(crimes_2012_joint))
```

**Question 7 (3 points):** _Next, make a table of *crime rates*. To do so, create a summary table showing counts of each type of offense by precinct in these data using group_by and summarise, grouping by precinct number, population, and offense type. Spread the data to wide format with the Offense types as column names and the count of reports as observations. Use kable to show the head of this table when done._

```{r}
#crime_by_loc<-st_as_sf(spokane_crimes_all,
 #                             coords=c(x = "Long", y = "Lat"), 
  #                            crs=4326,
   #                           remove=FALSE)
 # st_join(precincts_spok)

c_rates<-crimes_2012_joint%>%
  group_by(Precinct, Offense, totpop)%>%
 # summarise(c_rates, offense_count = count("Offense")/totpop)%>% #tricky line to write...
  summarise(count=n())
c_rates_wide<-pivot_wider(c_rates, names_from = Offense, 
              values_from = count,
              values_fill = 0)
 
  
kable(head(c_rates_wide))
```

**Question 8 (2 points)** _Select just the precinct number, population, and burglary rate from the precincts data frame. Then calculate the *burglary rate*, dividing the crimes by the population and multiplying by 100,000. Create a histogram using ggplot showing the distribution of this new variable._ 

```{r}
burg_rate<-c_rates_wide%>%
  select(Precinct, totpop, Burglary)%>%
  summarise(Burg_Rate = (Burglary/totpop)*100000)

ggplot(burg_rate, aes(x = Burg_Rate))+
  geom_histogram()+
  xlab("Burglary")+
  ylab("Count")
```

**Question 9 (2 points)** _Use left_join to join the precinct boundaries and burglary rate data frame. Then use tmap to create a map of the burglary rate. Use style="jenks" in the tm_polygons function to use natural breaks classifciation. What's one geographic trend you see on this map?_

```{r}
burg_rate_join<-precincts_spok%>%
  left_join(burg_rate)

tm_shape(burg_rate_join)+
  tm_polygons("Burg_Rate", style = "jenks")
```

{The areas from the center up thru the northern part of the city appear to have more burglaries than the south-side of the the city. The center of the city appears to have the worst rates of burglary.}

**Question 10 (3 points)** _The burglary rate for the city of Spokane in 2012, according to this dataset, is 1,657 per 100,000 residents. Using this figure, calculate the location quotient for burglary in each precinct within the city. Which precincts have the highest and lowest location quotients? What does this tell you about each of those precincts?_

```{r}
burg_lq<-burg_rate_join%>%
  mutate(burg_lq = Burg_Rate/1657)

kable(head(burg_lq))
#LOCATION QUOTIENT CALCULATION
```

{The location quotient tells us how the crime rate that corresponds to a particular precinct area compares to the overall burglary rate for the city. Precincts of particularly high traffic for an offense can be analyzed more closely to see why that may be.}

###Part 3: Kernel density analysis###
The last section of the lab asks you to create a kernel density map of burglaries.

**Question 11 (2 points)** _Create a new data frame from the original crime reports with just reported burglaries. Then create a ppp version of these data in the spatstat package using the methods shown in class. Plot this object to show the points._ 

```{r}
burg_only<-spokane_crimes_all%>%
  filter(Offense == "Burglary")

minlong<-min(burg_only$Long)
minlat<-min(burg_only$Lat)
maxlong<-max(burg_only$Long)
maxlat<-max(burg_only$Lat)

burg_ppp<-ppp(burg_only$Long,burg_only$Lat,c(minlong,maxlong),c(minlat,maxlat))

plot(burg_ppp)

```

**Question 12 (2 points)** _Create a kernel density map for burglaries in Spokane during the study period using spatstat using a kernel size that you believe best shows the distribution of reported crime._

```{r}
kernel_map<-density(burg_ppp)
plot(kernel_map)

plot(density(burg_ppp,0.008))
```
